---
title: "Homophily in Congress on Social Media"
author: "Marcel Julien Kalb, Mike Barth, Sam Känner"
format:
  pdf
---

<!-- # Instruction stuff -->

<!-- - Your report should be approximately 3 pages (approximately 1000 words excluding code and output). -->

<!-- - Make sure to include reference list, in particular for the theory part of the report. -->

# Introduction

For our analysis, we are concerning ourselves with the [Twitter Interaction Network for US Congress](https://snap.stanford.edu/data/congress-twitter.html) by the SNAP. It represents Twitter activity among members of the 117th U.S. Congress. The dataset contains 475 congressional Twitter accounts connected by 13,289 directed, weighted edges. An edge represents an observed interaction between two members, including retweets, quote tweets, replies, and mentions. Edge weights capture the relative strength of these interactions and are modeled as a normalized measure of influence between accounts. As the dataset itself does not contain any node properties besides the names and usernames of the members, we are enriching it using data from [voteview](https://voteview.com/congress/senate/117/text) and [github/unitedstates](https://github.com/unitedstates/congress-legislators) which cover chamber and party affiliation, age, as well as ideological position scores.

Our focus of analysis will be to assess homophily within the interaction network, based on different characteristics such as party affiliation, chamber and age. It has been shown quite thoroughly that party homophily exists, but we are most interested in whether tertiary factors such as chamber and age moderate this effect, e.g. do members of the senate exhibit more homophily than those in the house, and, do older members exhibit more or less homophily than their younger colleagues?

# Theory

Homophily is the principle that contact between similar people occurs at a higher rate than among dissimilar people. It can be further distinguished in two types: *status homophily* as the similarity of persons in their formal, informal or ascribed status and *value homophily* based on shared believes, values and attitudes of persons [(McPherson et al., 2001)](https://doi.org/10.1146/annurev.soc.27.1.415). 
We take the party affiliation, the chamber and the age as status homophily and expect that the number of ties between members of congress is shaped by their status, e.g. congress members mostly interact with members of their own party.
Age is expected to further shape the number of connections, since people tend to form ties with those of similar life stage [(McPherson et al., 2001)](https://doi.org/10.1146/annurev.soc.27.1.415), older legislators may have formed cross-party relationships before the current era of polarization, but also older members may tend to be less active on twitter. Also a legislators chamber structures the opportunity for interaction, as Senate and House members operate in largely separate institutional environments with different rules, schedules, and legislative responsibilities, making cross-chamber ties less likely to form regardless of party.
Value homophily is captured as left-right ideological position and stance on social and racial issues. There is a considerable tendency for adults to associate with those of their own political orientations [(McPherson et al., 2001)](https://doi.org/10.1146/annurev.soc.27.1.415). Therefore, we expect to find more interactions between members with similar positioning in the ideological dimensions. 



# Data Pre-Processing

```{r, echo=FALSE}
pacman::p_load("igraph", "dplyr", "purrr", "tibble", "stringr")

#### DATA ####

data <- jsonlite::fromJSON("congress_network/congress_network_data.json", simplifyVector = FALSE)
voteview <- readr::read_csv("HS117_members.csv", show_col_types=FALSE)

leg <- yaml::read_yaml(
  "https://raw.githubusercontent.com/unitedstates/congress-legislators/main/legislators-current.yaml",
  handlers = list(int = function(x) x)
)

soc <- yaml::read_yaml(
  "https://raw.githubusercontent.com/unitedstates/congress-legislators/main/legislators-social-media.yaml",
  handlers = list(int = function(x) x)
)

inList <- data[[1]]$inList
inWeight <- data[[1]]$inWeight
outList <- data[[1]]$outList
outWeight <- data[[1]]$outWeight
usernameList <- data[[1]]$usernameList

edges <- do.call(
  rbind, lapply(seq_along(outList), function(i) {
    if (length(outList[[i]]) == 0) return(NULL)
    data.frame(
      from = i - 1,                    
      to = unlist(outList[[i]]),
      weight = unlist(outWeight[[i]]))}))

g <- graph_from_data_frame(d = edges, directed = TRUE, vertices = data.frame(name = seq_along(usernameList) - 1, username = unlist(usernameList)))

#### Attributes ####

# V(g)$username # username
# V(g)$name # id

`%||%` <- function(x, y) if (is.null(x)) y else x

members <- purrr::map_dfr(leg, function(x) {
  terms <- x$terms
  
  purrr::map_dfr(terms, function(term) {
    tibble::tibble(
      bioguide_id = x$id$bioguide %||% NA_character_,
      first_name  = x$name$first %||% NA_character_,
      last_name   = x$name$last %||% NA_character_,
      full_name   = paste(x$name$first %||% "", x$name$last %||% ""),
      type        = term$type %||% NA_character_,
      state       = term$state %||% NA_character_,
      district    = as.character(term$district %||% NA_character_),
      party       = term$party %||% NA_character_,
      term_start  = as.Date(term$start),
      term_end    = as.Date(term$end)
    )
  })
})

social <- map_dfr(soc, function(x) {
  tibble(
    bioguide_id = x$id$bioguide %||% NA_character_,
    twitter = x$social$twitter %||% NA_character_
  )
})

members <- members %>%
  left_join(social, by = "bioguide_id")

members_117 <- members %>%
  filter(term_start < as.Date("2023-01-03"), term_end > as.Date("2021-01-03")) %>%
  mutate(
    chamber = case_when(
      type == "sen" ~ "Senate",
      type == "rep" ~ "House",
      TRUE ~ NA_character_
    ),
    twitter_clean = str_to_lower(str_remove(twitter, "^@"))
  )

members_117_one <- members_117 %>%
  filter(!is.na(twitter_clean), twitter_clean != "") %>%
  arrange(twitter_clean, term_start) %>%
  group_by(twitter_clean) %>%
  summarise(
    bioguide_id = first(bioguide_id),
    full_name = first(full_name),
    chamber = first(chamber),
    party = first(party),
    state = first(state),
    district = first(district),
    term_start = min(term_start, na.rm = TRUE),
    term_end = max(term_end, na.rm = TRUE),
    .groups = "drop"
  )

graph_vertices <- tibble(
  name = V(g)$name,
  username = V(g)$username,
  twitter_clean = str_to_lower(str_remove(V(g)$username, "^@")))

vertex_metadata <- graph_vertices %>%
  left_join(members_117_one, by = "twitter_clean")

V(g)$full_name <- vertex_metadata$full_name
V(g)$chamber <- vertex_metadata$chamber
V(g)$party <- vertex_metadata$party
V(g)$state <- vertex_metadata$state
V(g)$district <- vertex_metadata$district
V(g)$bioguide_id <- vertex_metadata$bioguide_id

# IDEOLOGY
nodes <- tibble(idx = seq_len(vcount(g)), bioguide_id  = V(g)$bioguide_id) %>%
  left_join(voteview, by = "bioguide_id")

V(g)$icpsr <- nodes$icpsr
V(g)$nominate_dim1 <- nodes$nominate_dim1
V(g)$nominate_dim2 <- nodes$nominate_dim2
V(g)$born <- nodes$born
V(g)$age24 <- (2024 - nodes$born) 
```

# Exploratory Data Analysis

<!-- - Describe key characteristics of the network. -->

<!-- - Calculate relevant network statistics. -->

<!-- - Interpret the results rather than simply reporting them. -->

<!-- - Connect findings back to the theoretical discussion. -->

Overall, the network is moderately sized and active, but not densely connected. Members form many retweet ties in absolute terms, yet these ties represent only a small share of all possible member-to-member connections ($\approx 5\%$). The the typical congress member has around 22–24 connections, while a small number of members have far higher connectivity. This suggests a selective and uneven retweet structure, with activity concentrated around more central or more active members rather than spread uniformly across the network. This is to be expected given in politics there are often central figures that may receive more attention, such as presidents or speakers.

```{r}
# Core descriptives: graph size

n_vertices <- vcount(g)
n_edges    <- ecount(g)
density    <- edge_density(g)

size_summary <- tibble::tibble(
  Measure = c(
    "Number of vertices",
    "Number of edges",
    "Edge density"
  ),
  Value = c(
    n_vertices,
    n_edges,
    density
  )
) %>%
  dplyr::mutate(Value = round(Value, 3))

cat(
  knitr::kable(
    size_summary,
    format = "pipe",
    col.names = c("Measure", "Value")
  ),
  sep = "\n"
)
```

```{r}
# Degree and weighted degree / strength

deg_in  <- degree(g, mode = "in")
deg_out <- degree(g, mode = "out")

str_in  <- strength(g, mode = "in")
str_out <- strength(g, mode = "out")

degree_summary <- tibble::tibble(
  Measure = c(
    "In-degree",
    "Out-degree",
    "Weighted in-strength",
    "Weighted out-strength"
  ),
  Min = c(
    min(deg_in, na.rm = TRUE),
    min(deg_out, na.rm = TRUE),
    min(str_in, na.rm = TRUE),
    min(str_out, na.rm = TRUE)
  ),
  Q1 = c(
    quantile(deg_in, 0.25, na.rm = TRUE),
    quantile(deg_out, 0.25, na.rm = TRUE),
    quantile(str_in, 0.25, na.rm = TRUE),
    quantile(str_out, 0.25, na.rm = TRUE)
  ),
  Median = c(
    median(deg_in, na.rm = TRUE),
    median(deg_out, na.rm = TRUE),
    median(str_in, na.rm = TRUE),
    median(str_out, na.rm = TRUE)
  ),
  Mean = c(
    mean(deg_in, na.rm = TRUE),
    mean(deg_out, na.rm = TRUE),
    mean(str_in, na.rm = TRUE),
    mean(str_out, na.rm = TRUE)
  ),
  Q3 = c(
    quantile(deg_in, 0.75, na.rm = TRUE),
    quantile(deg_out, 0.75, na.rm = TRUE),
    quantile(str_in, 0.75, na.rm = TRUE),
    quantile(str_out, 0.75, na.rm = TRUE)
  ),
  Max = c(
    max(deg_in, na.rm = TRUE),
    max(deg_out, na.rm = TRUE),
    max(str_in, na.rm = TRUE),
    max(str_out, na.rm = TRUE)
  )
) %>%
  dplyr::mutate(
    dplyr::across(where(is.numeric), ~ round(.x, 3))
  )

cat(
  knitr::kable(
    degree_summary,
    format = "pipe"
  ),
  sep = "\n"
)
```

```{r}
# Degree distributions

par(mfrow = c(1, 2))

hist(
  deg_in,
  main = "In-degree distribution",
  xlab = "In-degree"
)

hist(
  deg_out,
  main = "Out-degree distribution",
  xlab = "Out-degree"
)

par(mfrow = c(1, 1))
```

```{r}
# Structural network measures

recip <- reciprocity(g)

trans_global <- transitivity(g, type = "global")
trans_avg    <- transitivity(g, type = "average")

assort_degree <- assortativity_degree(g, directed = TRUE)

structure_summary <- tibble::tibble(
  Measure = c(
    "Reciprocity",
    "Global transitivity",
    "Average local transitivity",
    "Degree assortativity"
  ),
  Value = c(
    recip,
    trans_global,
    trans_avg,
    assort_degree
  )
) %>%
  dplyr::mutate(Value = round(Value, 3))

cat(
  knitr::kable(
    structure_summary,
    format = "pipe",
    col.names = c("Measure", "Value")
  ),
  sep = "\n"
)
```

The network shows a moderate level of reciprocity: about 46% of directed ties are reciprocated. This suggests that interactions are not purely one-sided, but often occur between pairs of members who engage with each other, possibly congress members working on the same topics or from similar geographic regions. Transitivity is also moderate, with global and average local clustering around 0.27–0.30, indicating that interaction patterns are somewhat clustered within groups of mutually connected members. This is to be expected given we are not accounting for chamber affiliation and or party affiliation at the moment, so clustering is to be expected. The slightly negative degree assortativity (-0.096) suggests that members with many outgoing interactions are not especially likely to direct those interactions toward members with many incoming interactions. In other words, highly active congress members are not disproportionately interacting with highly visible or frequently targeted congress members.

```{r}
# Homophily by party
el <- igraph::as_data_frame(g, what = "edges")
el$from_party <- V(g)$party[match(el$from, V(g)$name)]
el$to_party   <- V(g)$party[match(el$to, V(g)$name)]
mix <- xtabs(weight ~ from_party + to_party, data = el)
mix_prop <- mix / rowSums(mix)
cat("Overall\n", knitr::kable(round(mix_prop, 3), format = "pipe"), sep = "\n")
```

We can see very strong party homophily: Democratic members mostly interact with Democrats (0.878), and Republican members mostly interact with Republicans (0.881). Cross-party interaction is much lower, around 0.12 in both directions. The "Independent" category is too small to interpret. As we can see below, there does not seem to be much difference in these homophily structures between the house and senate.

```{r}
# Homophily by party by chamber
el$from_chamber <- V(g)$chamber[match(el$from, V(g)$name)]

for (ch in c("House", "Senate")) {
  el_ch <- el[!is.na(el$from_chamber) & el$from_chamber == ch, ]
  mix   <- xtabs(weight ~ from_party + to_party, data = el_ch)
  mix_prop <- mix / rowSums(mix)
  cat(ch, "\n", knitr::kable(round(mix_prop, 3), format = "pipe"), sep = "\n", "\n")
}
```

Left-right ideology is moderately strong (0.447), meaning members tend to interact with ideologically similar members. Assortativity on the social/racial dimension is weaker (0.119), and age assortativity is essentially negligible (0.038) which means that congress members do not interact more often with members of their age.

```{r}
# assortativity
leftrightass <- assortativity(g, V(g)$nominate_dim1, directed = TRUE) # L/R
socialracialass <- assortativity(g, V(g)$nominate_dim2, directed = TRUE) # social/racial issues
ageass <- assortativity(g, V(g)$age24, directed = TRUE) # age -> basically none

assort_table <- data.frame( dimension = c("Left-right", "Social/racial issues", "Age"), assortativity = c(leftrightass, socialracialass, ageass))

cat(
  knitr::kable(
    assort_table,
    digits = 3,
    col.names = c("Dimension", "Assortativity"),
    format = "pipe"
  ),
  sep = "\n"
)
```

```{r}
# Homophily by age
el$same_party <- !is.na(el$from_party) & 
  !is.na(el$to_party) & 
  el$from_party == el$to_party

# filter NAs
el_known <- el %>% filter(!is.na(from_party), !is.na(to_party))
# exclude Independents
el_known <- el_known[el_known$from_party != "Independent" & el_known$to_party != "Independent", ]

# per-node homophily score
homophily_score <- el_known %>%
  group_by(from) %>%
  summarise(total_w = sum(weight), ingroup_w = sum(weight[same_party]), h_score   = ingroup_w / total_w)

homophily_score$age <- V(g)$age24[match(homophily_score$from, V(g)$name)]
homophily_score$party <- V(g)$party[match(homophily_score$from, V(g)$name)]

# correlation
cat("Correlation between age and homophily:\n",cor(homophily_score$h_score, homophily_score$age, use = "complete.obs"), "\n\n")

# by party --> democrats weakly positive (+0.04), republicans kind of substantially negative (-0.128)
homophily_score %>%
  group_by(party) %>%
  summarise(`Correlation Age <-> Party Homophily` = cor(h_score, age, use = "complete.obs")) %>%
  mutate(`Correlation Age <-> Party Homophily` = round(`Correlation Age <-> Party Homophily`, 3)) %>%
  { cat(knitr::kable(., format = "pipe"), sep = "\n") }
```

Here, we are looking at whether the party homophily of individuals is higher or lower depending on their age. The correlation is very small and slightly negative (-0.034) overall, suggesting that older and younger members do not differ much in their tendency to interact within party. By party, the relationship is weakly positive for Democrats (0.048) and modestly negative for Republicans (-0.128), indicating that older Republican members are somewhat less party-homophilous in their tweet interactions.

# Visualisation

xxx

# Conclusion

xxx
